---
title: "Confidence in the Sample Mean"
author: "Chase Baggett"
date: "September 12, 2017"
output:
  github_document: default
  html_document: default
  pdf_document: default
---

###2.1 An investigator found that a random sample of size 100 from a normal population resulted in a 95 percent confidence interval for the population mean with length 4. The investigator would like a confidence interval with length about 1. About how many observations should the investigator take to obtain a confidence interval with length about 1?

Because the standard error is divided by the square root of the standard deviation, we say that information is gained at the square rate. Because of this, to make our confidence interval 4 times better, we have to increase our sample size by 16.

I have included an example in code. 
```{r}
perfect_normal <- function(n,mean,sd) { mean+sd*scale(rnorm(n)) }
original_sample_size <- 100
sample <- perfect_normal(100,mean=50,sd=40)
std.error <- function(x){sd(x)/sqrt(length(x))}
std.error(sample)

new_sample_size <- 1600
new_sample <- perfect_normal(new_sample_size,mean=50,sd=40)
std.error(new_sample)
```
###2.2 (b) The investigator would like to test simultaneously the hypotheses µ = 0 and σ^2 = 1 about µ and σ^2. Suggest a test statistic and a test procedure, and discuss their rationale.

```{r}
```

###2.3 It is reasonable to treat the data for problem 1.1 in your text as a random sample? If so, what is your reason for this conclusion and what is the population? If not, what is the point of studying the data?

No, it is not a random sample of the earth's population because they only collected data from UN countries. It might, however, be considered a random sample of the sub-population of UN member countries, as they don't explcitly tell us how they collected it.

Studying non-random samples can still be useful outside the realm of statistics. Without the desire to infer anything from the population, It still says something about the people who were sampled, even if we can't say anything about people not sampled. 

###2.4 Complete problem 1.1 in your text. In addition, print the graph requested as part of the solution to 1.1.3 and then sketch an estimate of the mean function by eye on the plot, describing the process that you used. Compare your estimate to a lowess smooth estimate of the mean function.

  1.1.1 The predictor is ppgdp in this case, and the response is fertility, because we want to know fertility's dependence on gdp per capita.
  1.1.2 A straght line does not appear to well characterize the relationship.
  1.1.3 However, for a log transformed x and y, a linear regression does seem plausible.
```{r}
library(alr4)
library(ggplot2)
library(gridExtra)
p1 <- ggplot(UN11,aes(x=ppgdp,y=fertility)) + geom_point(alpha=.5) + ggtitle("Fertility vs Per Capita GDP")
p2 <- ggplot(UN11,aes(x=log(ppgdp),y=log(fertility))) + geom_point(alpha=.5) + ggtitle("Naturtal Log of Fertility vs Natual Log of Per Capita GDP")
grid.arrange(p1,p2) 
```
###2.5 Assume that mheight in the data discussed in Section 1.1 in your text is a random sample from a normal distribution with mean µ and variance σ^2. 

####(a) derive the variance of the estimator ˆσ^2 of σ^2. 
####(b) Give the standard error of ˆσ^2.

###2.6 Assume that (mheight, dheight) in the data discussed in Section 1.1 in your text is a random sample from a bivariate normal population.
####(a) Estimate the correlation coefficient and describe what it is estimating.
####(b) Estimate the conditional mean E(mheight | dheight) and variance var(mheigh | dheight).
####(c) Estimate the conditional mean E(dheight | mheight) and variance var(dheigh | mheight), and describe the connection with the regression of part (b).
####(d) A mother’s height is 1 standard deviation (for mothers) above the average height for mothers. Her daughter’s height is expected to be ——— (fill in the blank with a number not a symbol) standard deviations (for daughters)